---
title: "Accessing the Wordbank database"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(wordbankr)
library(knitr)
library(dplyr)
library(ggplot2)
opts_chunk$set(message = FALSE, warning = FALSE, cache = FALSE)
theme_set(theme_bw())
```

The `wordbankr` package allows you to access data in the [Wordbank database](http://wordbank.stanford.edu/) from R. This vignette shows some examples of how to use the data loading functions and what the resulting data look like.


There are three different data views that you can pull out of Wordbank: by-administration, by-item, and administration-by-item.

The `get_administration_data` function gives by-administration information, for either a specific language and form or for all instruments:
```{r}
english_ws_admins <- get_administration_data("English", "WS")
head(english_ws_admins)
all_admins <- get_administration_data()
head(all_admins)
```

The `get_item_data` function gives by-item information, for either a specific language and form or for all instruments:
```{r}
spanish_wg_items <- get_item_data("Spanish", "WG")
head(spanish_wg_items)
all_items <- get_item_data()
head(all_items)
```

If you are only looking at total vocabulary size, `admins` is all you need, since it has both productive and receptive vocabulary sizes calculated. If you are looking at specific items or subsets of items, you need to load instrument data, using the `get_instrument_data` function. Pass it an instrument language and form, along with a list of items you want to extract (by `item_id`).
```{r}
eng_ws_canines <- get_instrument_data(instrument_language = "English",
                                      instrument_form = "WS",
                                      items = c("item_26", "item_46"))
head(eng_ws_canines)
```

By default `get_instrument_table` returns a data frame with columns of the administration's `data_id`, the item's `num_item_id` (numerical `item_id`), and the corresponding value. To include administration information, you can set the `administrations` argument to `TRUE`, or pass the result of `get_administration_data` as `administrations` (that way you can prevent the administration data from being loaded multiple times). Similarly, you can set the `iteminfo` argument to `TRUE`, or pass it result of `get_item_data`.

Loading the data is fast if you need only a handful of items, but the time scales about linearly with the number of items, and can get quite slow if you need many or all of them. So, it's a good idea to filter down to only the items you need before calling `get_instrument_data`.

As an example, let's say we want to look at the production of animal words on English Words & Sentences over age. First we get the items we want:
```{r, fig.width=6, fig.height=4}
animals <- get_item_data("English", "WS") %>%
  filter(category == "animals")
```

Then we get the instrument data for those items:
```{r}
animal_data <- get_instrument_data(instrument_language = "English",
                                   instrument_form = "WS",
                                   items = animals$item_id,
                                   administrations = english_ws_admins)
```

Finally, we calculate how many animals words each child produces and the median number of animals of each age bin:
```{r}
animal_summary <- animal_data %>%
  mutate(produces = value == "produces") %>%
  group_by(age, data_id) %>%
  summarise(num_animals = sum(produces, na.rm = TRUE)) %>%
  group_by(age) %>%
  summarise(median_num_animals = median(num_animals, na.rm = TRUE))
  
ggplot(animal_summary, aes(x = age, y = median_num_animals)) +
  geom_point()
```