---
title: "Accessing the Wordbank database"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(wordbankr)
library(knitr)
library(dplyr)
library(ggplot2)
opts_chunk$set(message = FALSE, warning = FALSE, cache = FALSE)
theme_set(theme_bw())
```

The `wordbankr` package allows you to access data in the [Wordbank database](http://wordbank.stanford.edu/) from R. This vignette shows some examples of how to use the data loading functions and what the resulting data look like.


## Common tables

Wordbank's `common` tables store data that's shared between all of the CDI instruments -- information about children, administrations, sources, items (questions), item categories, etc.

The `get_common_tables` function loads connections to all common tables:
```{r}
src <- connect_to_wordbank()
common_tables <- get_common_tables(src)
names(common_tables)
```

You generally don't need to access them directly, though. There are helper functions that load sensible groups of tables, process them, and join them together to give you information about each administration and item.

The `get_administration_data` function gives by-administration data:
```{r}
admins <- get_administration_data()
head(admins)
```

And the `get_item_data` function gives by-item data:
```{r}
items <- get_item_data()
head(items)
```

If you are only looking at total vocabulary size, `admins` is all you need, since it has both productive and receptive vocabulary sizes cached. If you are looking at specific items or subsets of items, you need to load instrument data.

## Instrument tables

Wordbank's `instrument` tables store each instrument's actual data: The response of every administration to every item.

To connect to an instrument's table, use `get_instrument_table`, passing it the instrument's language and form. You can then pass the instrument's table to `get_instrument_data`, along with a list of items you want to extract (by `item_id`).
```{r}
eng_ws_data <- get_instrument_data(instrument_language = "English",
                                   instrument_form = "WS",
                                   items = c("item_1", "item_42"))
head(eng_ws_data)
```

By default `get_instrument_table` just gets the `data_id` of the administration, the `num_item_id` (numerical `item_id`) of the item, and the corresponding value. To include administration data, you can set the `administrations` argument to `TRUE`, or pass the result of `get_administration_data` as `administrations` (that way you can prevent the administration data from being loaded multiple times). Similarly, you can set the `iteminfo` argument to `TRUE`, or pass it result of `get_item_data`.

Loading the data is fast if you need only a handful of items, but the time scales about linearly with the number of items, and can get quite slow if you need many or all of them. So, it's a good idea to filter down to only the items you need before calling `get_instrument_data`.

As an example, let's say we want to look at the production of animal words on English Words & Sentences over age. First we filter the items down to the ones we want:
```{r, fig.width=6, fig.height=4}
animals <- filter(items, language == "English", form == "WS",
                  category == "animals")

animal_data <- get_instrument_data(instrument_language = "English",
                                   instrument_form = "WS",
                                   items = animals$item_id,
                                   administrations = admins)

animal_summary <- animal_data %>%
  mutate(produces = value == "produces") %>%
  group_by(age, data_id) %>%
  summarise(num_animals = sum(produces, na.rm = TRUE)) %>%
  group_by(age) %>%
  summarise(median_num_animals = median(num_animals, na.rm = TRUE))
  
ggplot(animal_summary, aes(x = age, y = median_num_animals)) +
  geom_point()
```